WSClean: an implementation of a fast, generic wide-field imager for radio astronomy

### 2014-03-29 articles/2014/Article-WSClean/wsclean.tex
Emils comments

In this article, we present a new implementation of a generic wide-field imager that is significantly faster than \textsc{casa}'s $w$-projection implementation, and is able to successfully image any MWA observation. To obtain the increase in speed, the implementation uses the $w$-stacking method for correcting the $w$-terms, optionally combined with a new technique for $w$-snapshot imaging. We named the new imager ``WSClean'', as an abbreviation for ``$w$-Stacking Clean''. Our new imaging implementation, which we believe to be the fastest generic wide-field imager available, is publicly released\footnote{The \textsc{wsclean} source code can be found at:\\\href{http://sourceforge.net/p/wsclean}{http://sourceforge.net/p/wsclean}}.

@@

Imaging the full FOV with the full SKA becomes very expensive due to the high frequency and time resolution, and image resolution of $7.2\textrm{k} \times 7.2$k pixels. This translates to a computing power requirement of $\sim60$ PetaFLOPS.

\section{Conclusions} \label{sec:conclusions}
We have shown that the $w$-stacking algorithm is well-suited for imaging MWA observations. The \textsc{wsclean} $w$-stacking implementation is faster than \textsc{casa}'s $w$-projection algorithm in all common MWA imaging configurations, giving up to an order of magnitude increase in speed at a relatively small ZA of $10$\degree, and is slightly more accurate. Roughly speaking, for ZAs $>15\degree$ or FOVs $>35\degree$ snapshot imaging becomes faster. This can lead to performance improvements of a factor of $3$ for the MWA, but only in the most expensive imaging configurations that are less commonly used. Considering the extra regridding step required, which complicates issues such as calculating the integrated beam shape, recentering snapshots is worthwhile for the MWA only at very low elevations or with large fields of view. Extrapolation of the computing cost predicts that this holds for most arrays, including SKA low. A hybrid between $w$-stacking and $w$-projection does not improve performance over a pure $w$-stacking implementation, but does improve a pure $w$-projection implementation significantly. Our optimization of lowering the image resolution during inversion and prediction increases performance by a factor of 2--10 and has no noticeable effect on the accuracy.

@@

\textsc{wsclean} is currently not able to perform multi-scale clean. Results of applying \textsc{casa}'s multi-scale clean \citep{multiscale-clean-cornwell-2008} on MWA data show that especially the Galactic plane is significantly better deconvolved using multi-scale clean, but it is very expensive. Imaging a one-minute observation takes 30 hours of computational time without $w$-projection. We plan to implement some form of multi-scale cleaning in \textsc{wsclean}. It is likely that additional optimizations need to be made to be able to do this with acceptable performance. Combining multi-scale with wide-band deconvolution techniques is a possible further improvement \citep{rau-msmfs-2011}.

>>>>

In this article, we present a new implementation of a generic wide-field imager that is significantly faster than \textsc{casa}'s $w$-projection implementation. To obtain the increase in speed, the implementation uses the $w$-stacking method for correcting the $w$-terms, optionally combined with a new technique for $w$-snapshot imaging. We named the new imager ``WSClean'', as an abbreviation for ``$w$-Stacking Clean''. Our new imaging implementation, which we believe to be the fastest generic wide-field imager available, is publicly released\footnote{The \textsc{wsclean} source code can be found at:\\\href{http://sourceforge.net/p/wsclean}{http://sourceforge.net/p/wsclean}}.

@@

Imaging the full FOV with the full SKA becomes very expensive due to the high frequency and time resolution, and image size of $7.2\textrm{k} \times 7.2$k pixels. This translates to a computing power requirement of $\sim60$ PetaFLOPS.

\section{Conclusions} \label{sec:conclusions}
We have shown that the $w$-stacking algorithm is well-suited for imaging MWA observations. The \textsc{wsclean} $w$-stacking implementation is faster than \textsc{casa}'s $w$-projection algorithm in all common MWA imaging configurations, giving up to an order of magnitude increase in speed at a relatively small ZA of $10$\degree, and results in slightly lower imaging errors. Roughly speaking, for ZAs $>15\degree$ or FOVs $>35\degree$ snapshot imaging becomes faster. This can lead to performance improvements of a factor of $3$ for the MWA, but only in the most expensive imaging configurations that are less commonly used. Considering the extra regridding step required, which complicates issues such as calculating the integrated beam shape, recentering snapshots is worthwhile for the MWA only at very low elevations or with large fields of view. Extrapolation of the computing cost predicts that this holds for most arrays, including SKA low. A hybrid between $w$-stacking and $w$-projection does not improve performance over a pure $w$-stacking implementation, but does improve a pure $w$-projection implementation significantly. Our optimization of lowering the image resolution during inversion and prediction increases performance by a factor of 2--10 and has no noticeable effect on the accuracy.

@@

\textsc{wsclean} is currently not able to perform multi-scale clean. Results of applying \textsc{casa}'s multi-scale clean \citep{multiscale-clean-cornwell-2008} on MWA data show that especially the Galactic plane is significantly better deconvolved using multi-scale clean, but it is very computationally expensive. Imaging a one-minute observation takes 30 hours of computational time without $w$-projection. We plan to implement some form of multi-scale cleaning in \textsc{wsclean}. It is likely that additional optimizations need to be made to be able to do this with acceptable performance. Combining multi-scale with wide-band deconvolution techniques is a possible further improvement \citep{rau-msmfs-2011}.
